Football Championship Ticket Options

### 2018-10-10 paper.Rmd
updated code

top25InitialPrices = initialData %>% 
  slice(., 1L) %>% 
  dplyr::filter(masseyYearEnd < 26) %>% 
  ungroup() %>% 
  mutate(name = factor(name, levels = .$name)) %>% 
  select(masseyYearEnd, UL_price, LL_price, name) %>% 
  mutate(average_price = (as.numeric(UL_price) + as.numeric(LL_price)) / 2) %>% 
  tidyr::gather(type, value, -name, -masseyYearEnd) %>% 
  arrange(name) %>% 
  ggplot(., aes(name, as.numeric(value), color = type)) +
  geom_point(size = 3, alpha = .5) +
  theme_minimal() +
  scale_color_brewer(palette = "Dark2")

>>>>

top25InitialPrices = initialData %>% 
  slice(., 1L) %>% 
  dplyr::filter(masseyYearEnd < 26) %>% 
  ungroup() %>% 
  mutate(name = factor(name, levels = top25PrevNames$Team)) %>% 
  select(masseyYearEnd, UL_price, LL_price, name) %>% 
  mutate(average_price = (as.numeric(UL_price) + as.numeric(LL_price)) / 2) %>% 
  tidyr::gather(type, value, -name, -masseyYearEnd) %>% 
  arrange(name) %>% 
  ggplot(., aes(name, as.numeric(value), color = type)) +
  geom_point(size = 3, alpha = .5) +
  theme_minimal() +
  scale_color_brewer(palette = "Dark2")

### 2018-10-10 paperFrozen.Rmd
updated code

initialData = readr::read_csv("data/optionsDat.csv")

previousYearRank = readr::read_csv("data/massey2017YearEnd.csv")

@@

completeData = readr::read_csv("data/completeData.csv") %>% 
  mutate(day = gsub("\\s.*", "", .$dateTime))

@@

allHistory = read.csv("data/allHistory.csv")
```

@@

Since its adoption in the 2014-2015 season, the College Football Playoff (CFP) has given college football fans even more reason to be excited about rankings. No longer did a team need to be in the top 2 of the Bowl Championship Series (BCS) to be a championship contender -- instead, the top 4 teams entered into a playoff. In capitalizing on the intersection between excitement and fandom, the CFP teamed with OptionIt to introduce an RSVP system for championship game tickets, starting in 2018. A fan may by an RSVP (an option) for a team. If that team would make it to the championship game, then the fan holding the option would be able to purchase a ticket at face value. In the end, this is likely to save fans a considerable amount of money as compared to purchasing tickets on the secondary market -- tickets on the secondary market were estimated to be an average of \$2,689 for the 2018 Championship game. In financial modeling terms, this system could be considered a call option for speculation. When buying an option (the RSVP), the buyer has the right to buy a stock (a ticket) in the future at the strike price (the face value of the ticket).

@@

When examining the option prices, it becomes apparent that the best teams have the highest prices; conversely teams with little chance of making the CFP have very low dollar values. Among those top teams, though, what might be driving the prices? As a preliminary pass, we are going to look at current values and volatility, and what impact wins, injuries, strength of schedule, and rankings have on those metrics. Specifically, we would anticipate that volatility has a positive relationship with wins (i.e., the more wins a team has, the more the option changes), but a negative relationship with rankings (the lower the ranking, the more the option changes). Number of injuries, while certainly challenging to a given team, might not be critical at this point within the season. Nonetheless, we would anticipate that the number of injuries will drive the option value down. We can certainly imagine a world in which strength of schedule has an impact on prices and volatility -- a nightmarish schedule might all but destroy your chances of making it to the playoffs.

@@

The internet is full of interesting data and finding it is never too hard. While finding interesting data does not present a challenge, finding data that is free to use can be tricky. In the context of data, free to use does not necessarily mean free in a monetary sense. Instead of a paywall, data is often locked behind terms of service (ToS). While scraping data has become easier, data creators are getting more sophisticated at "hiding" the data (i.e., it is not sitting nicely within the html, but loads through JavaScript) or ToS. With the appropriate acumen, the location of the data ceases to pose a challenge; the same cannot be said for ToS. In scraping our data, we not only read the ToS, but we also examined each site's robots.txt. Since we wanted to conduct our data collection in a programmatic fashion, we wanted to ensure that we were not going to be running afoul of restrictions placed within the robots.txt file. To that end, all the data that we collected is free to use -- in other words, no restrictions apply.

@@

In an effort to reduce visual clutter and focus on potentially high-value teams, we can visually inspect the top 25 teams at the end of the 2017-2018 season.

```{r, fig.height=3.5}
top25InitialPrices + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(), 
        legend.position = "bottom", 
        legend.title = element_blank())

@@

@@

The models presented here are incomplete, given that we are only 3 weeks into the college football season. Despite this, we are encouraged that the ticket options are behaving in a predictable manner. As data continues to roll in, our models will grow in complexity and predictive power. Ultimately, the goal is buy before a team becomes the hot dark horse and sell become a team's dreams burn in flames.

>>>>

initialData = readr::read_csv("data/frozen/optionsDat.csv")

previousYearRank = readr::read_csv("data/frozen/massey2017YearEnd.csv")

@@

completeData = readr::read_csv("data/frozen/completeData.csv") %>% 
  mutate(day = gsub("\\s.*", "", .$dateTime))

@@

allHistory = read.csv("data/frozen/allHistory.csv")
```

@@

Since its adoption in the 2014-2015 season, the College Football Playoff (CFP) has given college football fans even more reason to be excited about rankings. No longer did a team need to be in the top 2 of the Bowl Championship Series (BCS) to be a championship contender -- instead, the top 4 teams entered into a playoff. In capitalizing on the intersection between excitement and fandom, the CFP teamed with OptionIt to introduce an RSVP system for championship game tickets, starting in 2018. A fan may buy an RSVP (an option) for a team. If that team would make it to the championship game, then the fan holding the option would be able to purchase a ticket at face value. In the end, this is likely to save fans a considerable amount of money as compared to purchasing tickets on the secondary market -- tickets on the secondary market were estimated to be an average of \$2,689 for the 2018 Championship game. In financial modeling terms, this system could be considered a call option for speculation. When buying an option (the RSVP), the buyer has the right to buy a stock (a ticket) in the future at the strike price (the face value of the ticket).

@@

When examining the option prices, it becomes apparent that the best teams have the highest prices; conversely teams with little chance of making the CFP have very low dollar values. Among those top teams, though, what might be driving the prices? As a preliminary investigation, we are going to look at current values and volatility, and what impact wins, injuries, strength of schedule, and rankings have on those metrics. Specifically, we would anticipate that volatility has a positive relationship with wins (i.e., the more wins a team has, the more the option changes), but a negative relationship with rankings (the lower the ranking, the more the option changes). Number of injuries, while certainly challenging to a given team, might not be critical at this point within the season. Nonetheless, we would anticipate that the number of injuries will drive the option value down. We can certainly imagine a world in which strength of schedule has an impact on prices and volatility -- a nightmarish schedule might all but destroy your chances of making it to the playoffs.

@@

The internet is full of interesting data and finding it is never too hard. While finding interesting data does not present a challenge, finding data that is free to use can be tricky. In the context of data, free to use does not necessarily mean free in a monetary sense. Instead of a paywall, data is often locked behind terms of service (ToS). While scraping data has become easier, data creators are getting more sophisticated at "hiding" the data (i.e., it is not sitting nicely within the html, but loads through JavaScript) or limiting use through restrictive ToS. With the appropriate acumen, the location of the data ceases to pose a challenge; the same cannot be said for ToS. In scraping our data, we not only read the ToS, but we also examined each site's robots.txt. Since we wanted to conduct our data collection in a programmatic fashion, we wanted to ensure that we were not going to be running afoul of restrictions placed within the robots.txt file. To that end, all the data that we collected is free to use -- in other words, no restrictions apply as far as we know.

@@

In an effort to reduce visual clutter and focus on potentially high-value teams, we visually inspect the top 25 teams at the end of the 2017-2018 season.

```{r, fig.height=3.25}
top25InitialPrices + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        axis.title.x = element_blank(), 
        axis.title.y = element_blank(), 
        legend.position = "bottom", 
        legend.title = element_blank())

@@

$$ Daily_\Delta = \frac{price - price_{lag}}{price_{lag}} - 1 $$
$$ volatility = \sigma_{daily_\Delta} * \sqrt[]{252}$$

@@

The models presented here are incomplete, given that we are only 3 weeks into the college football season. Despite this, we are encouraged that the ticket options are behaving in a predictable manner. As data continues to roll in, our models will grow in complexity and predictive power. Ultimately, the goal is buy before a team becomes the hot dark horse and sell before a team's dreams burn in flames.
